Towards a standardized evaluation of multiple imputation routines

### 2022-03-05 Manuscript/StandardizedEvaluation.Rmd
Edit intro and problems section

---
title: "Towards a standardized evaluation of multiple imputation routines"
author: "Hanne Oberman and Gerko Vink"
date: "19-1-2022"
abstract: "Developing new imputation methodology has become a very active field. Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods. In this paper I propose a move towards a standardized evaluation of imputation methods. To demonstrate the need for standardization, I highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines. This may lead to suboptimal use of multiple imputation in practice. Additionally, I suggest a course of action for simulating and evaluating missing data problems."
bibliography: StandardizedEvaluation.bib  
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
source(here::here("R/utils.R"))
source(here::here("R/run_sim.R"))
```

# TO DO

- add ADEMP paper [@morr18], diagnostic checking paper [@zhao22]
- add PPC (refer to Mingyang, and Gelman & King)
- expand convergence section
- add RMSE
- add insights from literature review for chapter on missing data in EHR
- add link to simulation script, and $n$
- add simulation results

@@

Multiple imputation @rubin1987 is a state-of-the-art technique for drawing valid conclusions from incomplete data, and is increasingly used in many substantive fields, such as medicine [see e.g. @pulkki2015cumulative; @alkire2015global; @stinesen2015reduced], psychology [e.g. @sorbi2014medium], economics [e.g. @wong2015economic; @dafoe2013democratic; @allan2015determinants], sociology [e.g. @crowley2014flexible; @de2014religiosity], and official statistics [e.g. @little2012calibrated; @klausch2015selection].

The technique has earned a permanent spot in research and policy making, demonstrated e.g. by the detailed manual created by the National Research Council [@little2012prevention]. Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing multiple imputation techniques. After all, the principle of multiple imputation is very intuitive.

The idea behind multiple imputation is to impute (fill in) the missing values multiple times to obtain a valid estimate of what could have been. The multiple datasets that are thus obtained can be analyzed by standard techniques and the analysis results can be combined into a single inference. In contrast to ad hoc methods for dealing with missing values (e.g. list-wise deletion, mean imputation, regression imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.

The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values. In general it holds that modeling missing data becomes more challenging when the amount of missingness increases. However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness. In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion.

When evaluating the statistical properties (and thereby the practical applicability) of new imputation methodology, researchers most often make use of simulation studies. In such studies a complete data set is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of the imputation method. However, no golden standard has been established and, as a result, validity of the simulation and evaluation procedures may differ tremendously from one developer to another.

The purpose of this paper is threefold: First, I raise some concerns with respect to evaluating imputation methodology. These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals. Second, I provide imputation methodologists with a suggested course of action when simulating missing data problems. This suggested approach should identify a common ground, but is in no way intended as an absolute solution. This identifies the third purpose of this paper: discussion. I hope to elicit critical thinking with respect to the problems at hand. We are all convinced that our methodology has some merit. But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \emph{proving the point} and would legitimately consider the statistical properties.

## Why some evaluations should not be trusted

As of today, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods. Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers. This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to suboptimal use of multiple imputation in practice. To demonstrate the broad impact of these problems, I subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation. I further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.

@@

To evaluate the ability of an imputation routine to handle the non-response model, a form of truth has to be established. Those who perform simulation studies are in the luxury position to establish the truth beforehand, either by generating data from a statistical model, or by sampling data from a sufficiently large observed set. However, when generating data, two easily overlooked problems may arise.

First, data are often generated following the model (or distribution) that is also used for imputing the data. The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied. Although such procedures may be statistically relevant, the approach would be no good for even the simplest case. Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution.

Second, data are often generated such that the problem that is being studied is most pronounced. This results in simulated data that contains very valuable information structures, i.e. the correlations between groups of variables may be very pronounced. In other words, no matter what type of missingness is induced, the observed parts of the data may still hold much, if not all, of the information about the missing part and the performance of the imputation procedure is not surprisingly evaluated as very good. Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.

```{r sim_nobs, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, paged.print=TRUE}
# setup
set.seed(1)
n_sim <- 100

sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
  purrr::map_dfr(c(50, 500, 5000), ~{
    run(n_obs = .x) %>% cbind(sim = .i, .)})})

# gtsummary::tbl_summary(sim_nobs, by = dat)
# https://haozhu233.github.io/kableExtra/awesome_table_in_html.html#Group_rows_via_labeling
sim_nobs %>% 
  group_by(n_obs, dat) %>% 
  summarise(
    mean_Y = mean(ave),
    beta_X = mean(est),
    CIW = mean(ciw),
    CR = mean(cov),
    RMSE_pred = mean(rmse_pred),
    RMSE_cell = mean(rmse_cell, na.rm = TRUE))  %>% 
  .[, -1] %>% 
  knitr::kable(
    col.names = c("", "$\\mu_{Y} = 10$", "$\\beta_{X} = 0.4$", "CIW", "CR", "RMSE$_{p}$", "RMSE$_{c}$"), 
    digits = 3,
    escape = FALSE) %>% 
  kableExtra::kable_classic(full_width = TRUE) %>% 
  kableExtra::pack_rows(group_label = "$n = 50$", 1, 3) %>% 
  kableExtra::pack_rows(group_label = "$n = 500$", 4, 6) %>% 
  kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9)

@@

In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94. The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR). Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g. @rubi76, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.

@@

Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit @molenberghs2008every - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios. In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis see e.g. @molenberghs2014handbook, part 5.

@@

The amount of missingness that is used to simulate the performance of imputation methods differs between studies. Some studies use only 10% missing data where other studies push the limits to additionally investigate the performance under at least 50% missing data. This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions. This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.

@@

The evaluation criteria used to assess imputation performance vary from one developer to another. This is not surprising as people from different fields could have a different focus on the problem at hand. However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated. I highlight four aspects of evaluation that should be considered when evaluation imputation routines.

First, developers often only inquire about the 'accuracy' (i.e. how well can the method reproduce the original data). The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied. After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p. 591.

Second, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny. Rather, one would prefer an imputation technique to yield both valid inference and plausible imputations. It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions.

Third, sampling variation has always been an essential part of the evaluation of multiple imputation methodology. However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink14. After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.

@@

@@

For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches. The parameters of that single complete set will serve as comparative truth. This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations. One needs to realize, however, that the conventional pooling rules cf. @rubin1987, pp. 76-77, do not apply for finite population inference and that alternative pooling rules need to be used @raghunathan2003multiple, @vink2014pooling.

@@

Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching @little1988missing, @rubin1986statistical may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

- \emph{Bias:} Results should preferable be unbiased. However, the way bias is considered can greatly influence the interpretation of the results. For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered. The way bias is computed should therefore be carefully chosen and described. 
- \emph{Interval width:} The way the confidence interval is calculated should be described. Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore. 
- \emph{Coverage:} Coverage of a 95% interval should in theory be $\geq 95$, where a coverage of 95% would be most efficient. Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained. 
- \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data. Under anything but the MCAR assumption, this can be expected. When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed see @abayomi2008diagnostics for an detailed overview of diagnostic evaluation for multivariate imputations. When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study. 
- \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference. However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property. When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values. 
- \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers, PIGS, @li2012imputing. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.

>>>>

---
title: "Towards a standardized evaluation of multiple imputation routines"
author: "Hanne Oberman and Gerko Vink"
date: "19-1-2022"
abstract: "Developing new imputation methodology has become a very active field. Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods. In this paper we propose a move towards a standardized evaluation of imputation methods. To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines. This may lead to suboptimal use of multiple imputation in practice. Additionally, we suggest a course of action for simulating and evaluating missing data problems."
bibliography: StandardizedEvaluation.bib  
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
library(dplyr)
# source(here::here("R/utils.R"))
# source(here::here("R/run_sim.R"))
```

<!-- # Structure -->

<!-- 1. introduction -->

<!-- - aim -->
<!-- - number of simulations -->

<!-- 2. data generation -->

<!-- - model vs design based -->
<!-- - data types -->
<!-- - sample vs population (vs bootstrapping??) -->
<!-- - number of observations -->
<!-- - number of variables -->
<!-- - coherence between variables -->

<!-- 3. missingness generation -->

<!-- - monotone/univariate vs multivariate -->
<!-- - missingness mechanism -->
<!-- - type of M(N)AR -->
<!-- - missingness proportion (vs proportion of incomplete cases! 50% incomplete == 25% missing in bivariate data) -->

<!-- 4. missing data methods -->

<!-- - imputation method -->
<!-- - number of imputations -->
<!-- - number of iterations -->

<!-- 5. performance evaluation -->

<!-- - estimand -->
<!-- - diagnostic -->
<!-- - ppc -->

@@

Multiple imputation [@rubin1987] is a state-of-the-art technique for drawing valid conclusions from incomplete data. The technique has earned a permanent spot in research and policy making, demonstrated e.g. by the detailed manual created by the National Research Council [@little2012prevention]. Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing multiple imputation techniques. After all, the principle of multiple imputation is very intuitive.

The idea behind multiple imputation is to impute (fill in) the missing values in incomplete data multiple times, to obtain a valid estimate of what could have been observed. The multiple datasets that are thus obtained can be analyzed by standard techniques and the analysis results can be combined into a single inference. In contrast to ad hoc methods for dealing with missing values (e.g. list-wise deletion, mean imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account. [It has been empirically demonstrated time and again that the inferences on multiply imputed data yield unbiased and confidence valid results (through Rubin's rules). But how do we know that this is the case, if we do not have access to the true but unobserved values of the missing entries? And how can we be sure that extensions to the methodology yield valid inferences too? Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods. That's why we use simulation studies for the evaluation of imputation routines.]

<!-- The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values. In general it holds that modeling missing data becomes more challenging when the amount of missingness increases. However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness. In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion. -->

When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies. In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, [then several missing data methods are applied] and a set of evaluation criteria is postulated to evaluate the performance of the imputation methods. However, no golden standard has been established to evaluate these imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.

The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology. These concerns stem from careful consideration with fellow 'imputers' and from encounters as a reviewer for statistical journals. Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems. This suggested approach should identify a common ground, but is in no way intended as an absolute solution. This identifies the third purpose of this paper: discussion. We hope to elicit critical thinking with respect to the problems at hand. We are all convinced that our methodology has some merit. But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \emph{proving the point} and would legitimately consider the statistical properties.

# Why some evaluations should not be trusted

As of today, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods. Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers. This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to sub-optimal use of multiple imputation in practice. To demonstrate the broad impact of these problems, we subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation. we further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.

@@

To evaluate the ability of an imputation routine to handle ~~the non-response model~~ [missing data entries], a form of truth has to be established. Those who perform simulation studies are in the luxury position to establish the truth beforehand, by choosing a data generating mechanism to obtain a complete dataset.  Complete data is typically either generated from a statistical model, or sampled from a sufficiently large ~~observed~~ set [model-based vs design-based simulation]. ~~However, when generating data, two easily overlooked problems may arise.~~

The problem with model-based data generating mechanisms is that method performance may not translate to empirical data. First, data are often generated following the model (or distribution) that is also used for imputing the data. The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied [@morr18]. Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case. Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution. [Add that data types may be more messy.] Second, data are often generated such that the problem that is being studied is most pronounced. This results in simulated data that contains very valuable information structures, i.e. the correlations between groups of variables may be very pronounced. In other words, no matter what type of missingness is induced, the observed parts of the data may still hold much, if not all, of the information about the missing part and the performance of the imputation procedure is not surprisingly evaluated as very good. Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.

The problem with design-based simulation is that it's difficult to find large datasets without missing entries. And imputing the missing data beforehand and then treating the once-imputed data as comparative truth may favor the used imputation method in further evaluations. A second issue is that SEs may be incorrect if the (re)sampling is not handled correctly. [Add pooling when the sample is the population [@vink14] and write something about bootstrapping SEs].

Problems with either of these data generation approaches is that the number of observations, the number of variables, the variable type(s) and the coherence between variables may give some imputation methods a comparative advantage, if they are not a) varied across simulation conditions [when is this fair? is this even an option?], or b) the object of the study [add some examples of simulations specifically focused on small samples, categorical data, etc.].

[Maybe add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable. I don't know where to introduce the difference between statistical inference and prediction. But this is an important third category, it seems. These studies apply different combinations of imputation and estimation methods on one or more benchmark datasets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods. In this context, there is no talk of retrieving the 'true' but unobserved values at all. Refer to missing data chapter [@liu21].]

```{r sim_nobs, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, paged.print=TRUE}
# # setup
# set.seed(1)
# n_sim <- 100
# 
# sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
#   purrr::map_dfr(c(50, 500, 5000), ~{
#     run(n_obs = .x) %>% cbind(sim = .i, .)})}) 
# 
# # gtsummary::tbl_summary(sim_nobs, by = dat)
# # https://haozhu233.github.io/kableExtra/awesome_table_in_html.html#Group_rows_via_labeling
# sim_nobs %>% 
#   group_by(n_obs, dat) %>% 
#   summarise(
#     mean_Y = mean(ave),
#     beta_X = mean(est),
#     CIW = mean(ciw),
#     CR = mean(cov),
#     RMSE_pred = mean(rmse_pred),
#     RMSE_cell = mean(rmse_cell, na.rm = TRUE))  %>% 
#   .[, -1] %>% 
#   knitr::kable(
#     col.names = c("", "$\\mu_{Y} = 10$", "$\\beta_{X} = 0.4$", "CIW", "CR", "RMSE$_{p}$", "RMSE$_{c}$"), 
#     digits = 3,
#     escape = FALSE) %>% 
#   kableExtra::kable_classic(full_width = TRUE) %>% 
#   kableExtra::pack_rows(group_label = "$n = 50$", 1, 3, escape = FALSE) %>% 
#   kableExtra::pack_rows(group_label = "$n = 500$", 4, 6, escape = FALSE) %>% 
#   kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9, escape = FALSE)
```

@@

Missingness comes in many shapes and forms [platitude, but how else to start?]. The most obvious characteristics of missing data are the missingness pattern(s) and missingness mechanism(s). Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data [@litt20, p. 8].

<!-- L&R, p. 8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".-->

The missing data pattern may inadvertently guide which imputation method prevails in simulation studies. Some ad hoc methods are known to yield valid inferences in very specific missing data patterns (e.g., list-wise deletion when there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case. Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions [re-read this! and add monotone/univariate vs multivariate and sporadic vs systematic dichotomies]. Related to the missingness pattern across variables is the missingness proportion. The terminology on this topic can be confusing. Does a missingness proportion of 50% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry? We will refer to the latter as the proportion of incomplete cases. But this is not a clear distinction in the literature, which may lead to incorrect conclusions [re-phrase this too]. Moreover, the amount of missingness that is used to simulate the performance of imputation methods differs between studies. [Add ref to chapter [@liu21] and add info.] Some studies use only 10% missing data where other studies push the limits to additionally investigate the performance under at least 50% missing data. This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions. This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.

[Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section). This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and sometimes realistic in practice, 3) that MAR can be generated *spuriously* if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice. And finally something about types of M(N)AR, but what are specific problems with that?]

# 
\hrule

#
[Until next hline: to be edited!] In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94. The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR). Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g. @rubi76, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.

@@

Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit [@molenberghs2008every] - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios. In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis see e.g. @molenberghs2014handbook, part 5.

@@

# 
\hrule

@@

[Add bridge from last section. And explain how this part is subdivided: first general characteristics of the imputations such as convergence, then performance measures.]

[Performance of different imputation methods may be influenced by preconditions of the methods themselves. All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with 100s of imputations. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.]

Moreover, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny. Rather, one would prefer an imputation technique to yield both valid inference and plausible imputations. It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions.

[Then onto the performance measure part.]

The evaluation criteria used to assess imputation performance vary from one developer to another. This is not surprising as people from different fields could have a different focus on the problem at hand. But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.

<!-- ~~However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated. We highlight four aspects of evaluation that should be considered when evaluation imputation routines.~~ -->

Developers often only inquire about the 'accuracy' (i.e. how well can the method reproduce the original data). [Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied. After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p. 591. [So, what could go wrong by focusing on accuracy exclusively?]

[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]

[Add problems with RMSE here.]

Sampling variation has always been an essential part of the evaluation of multiple imputation methodology. However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink14. After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.

@@

# Illustration

[Add simulation study + results here!]

@@

For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches. The parameters of that single complete set will serve as comparative truth. This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations. One needs to realize, however, that the conventional pooling rules [cf. @rubin1987, pp. 76-77], do not apply for finite population inference and that alternative pooling rules need to be used [@raghunathan2003multiple; @vink2014pooling].

@@

Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching [@little1988missing; @rubin1986statistical] may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

- \emph{Bias:} Results should preferable be unbiased. However, the way bias is considered can greatly influence the interpretation of the results. For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered. The way bias is computed should therefore be carefully chosen and described. 
- \emph{Interval width:} The way the confidence interval is calculated should be described. Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore. 
- \emph{Coverage:} Coverage of a 95% interval should in theory be $\geq 95$, where a coverage of 95% would be most efficient. Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
- [Add RMSE 2x]

- \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data. Under anything but the MCAR assumption, this can be expected. When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed see @abayomi2008diagnostics for an detailed overview of diagnostic evaluation for multivariate imputations. When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study. 
- \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference. However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property. When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values. [Add PPC]
- \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers, PIGS, @li2012imputing. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered. [Extend this part!]
- [Add PPC (Mingyang) [@nguy17], [@zhao22]]
